---
title: "run FinOMOP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{run}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
devtools::load_all(".")
```

```{r}

base_path <- r"(C:\Users\javier\REPOS\GITLAB_FinOMOP\finomop_omop_vocabulary)"

path_to_input_mapping_vocabularies_info_file <- file.path(base_path, "VOCABULARIES", "vocabularies_info.csv")
path_to_input_omop_vocabulary_folder <- file.path(base_path, "OMOP_VOCABULARIES", "input_omop_vocabulary")
path_to_input_database_counts_file <- file.path(base_path, "CODE_COUNTS", "databases_coverage.csv")
path_to_input_vocabularies_coverage_file <- file.path(base_path, "CODE_COUNTS", "vocabularies_coverage.csv")

path_to_temp_omop_vocabulary_folder <- file.path(base_path, "OMOP_VOCABULARIES", "temp_omop_vocabulary")

path_to_output_omop_vocabulary_folder <- file.path(base_path, "OMOP_VOCABULARIES", "output_omop_vocabulary")
path_to_output_dashboard_file <- file.path(base_path, "StatusReport", "dashboard.html")
```



# Read vocabularies_info.csv 

vocabularies_info.csv contains the information to build the vocabularies 

```{r}
mapping_tables <- importMappingTables(path_to_input_mapping_vocabularies_info_file)

usagi_mapping_tables <- mapping_tables$usagi_mapping_tables
vocabulary_info_mapping_tables <- mapping_tables$vocabulary_info_mapping_tables

usagi_mapping_tables
vocabulary_info_mapping_tables
```


temp fix:icd9fi, 
```{r}
usagi_mapping_tables[[4,5]][[1]] <- usagi_mapping_tables[[4,5]][[1]] |> 
dplyr::distinct(sourceCode, .keep_all = TRUE )
```


```{r}
usagi_mapping_tables <- usagi_mapping_tables |> validateTables("UsagiForCCR")
vocabulary_info_mapping_tables <- vocabulary_info_mapping_tables |> validateTables("VocabularyInfo")

usagi_mapping_tables |> filter(n_failed_rules!=0)
vocabulary_info_mapping_tables |> filter(n_failed_rules!=0)
```


# Convert mapping files to OMOP tables 

This will only run these with no errors
```{r}
convertMappingsTablesToOMOPtables(
  usagi_mapping_tables,
  vocabulary_info_mapping_tables,
  path_to_temp_omop_vocabulary_folder, 
  ignore_failed_rules = FALSE
)
```


# merge with ahtena vocabularies 

```{r}
mergeOMOPtables(path_to_input_omop_vocabulary_folder,
                path_to_temp_omop_vocabulary_folder,
                path_to_output_omop_vocabulary_folder)
```



## Validate the newely created vocabularies
```{r}
omop_tables <- importOMOPtables(path_to_output_omop_vocabulary_folder, c("CONCEPT","VOCABULARY", "CONCEPT_CLASS", "CONCEPT_RELATIONSHIP", "CONCEPT_SYNONYM"))

omop_tables <- validateTables(omop_tables)

omop_tables
```


## Import code counts

```{r}
databases_code_counts_tables <- importDatabasesCodeCountsTables(path_to_input_database_counts_file)

databases_code_counts_tables <- validateTables(databases_code_counts_tables, table_type = "CodeCounts")

databases_code_counts_tables

```

autofix
```{r}

databases_code_counts_tables <- databases_code_counts_tables |> 
  mutate(table = if_else(n_failed_rules==0, table, 
                                map2(.x=table, .y=name, .f=~autoFixDatabaseCodeCountTable(.x,.y))
                                ))

databases_code_counts_tables <- validateTables(databases_code_counts_tables, table_type = "CodeCounts")

databases_code_counts_tables

```



# check coverage 

```{r}
mapping_status <- calculateMappingStatus(
  path_to_input_vocabularies_coverage_file,
  omop_tables, 
  databases_code_counts_tables)
```


```{r}
plotTableMappingStatus(mapping_status)
```

```{r}
plotTableVocabularyStatus(mapping_status, "ICD10fi")
```

# Build report

```{r}
tmp_html <- path_to_output_dashboard_file
buildStatusDashboard(
  usagi_mapping_tables = usagi_mapping_tables,
  vocabulary_info_mapping_tables = vocabulary_info_mapping_tables,
  omop_tables  = omop_tables, 
  databases_code_counts_tables = databases_code_counts_tables, 
  mapping_status  = mapping_status,
  output_file_html = tmp_html)
browseURL(tmp_html) 
```



# Test against posgress

Start postgress database and rung pgAdmin4

```{bash} 
pg_ctl -D /c/Users/javier/WORKSPACES/OHDSI/postgres  -logfile start

/c/Users/javier/AppData/Local/postgresql-14.4-1-windows-x64-binaries/pgsql/pgAdmin\ 4/bin/pgAdmin4.exe 
```


password `5Yjyc%:R.0YmLh[pm<N%`

```{r}
# For those interested in seeing the CDM changes from 5.3 to 5.4, please see: http://ohdsi.github.io/CommonDataModel/cdm54Changes.html

cd_admin <- DatabaseConnector::createConnectionDetails(
  dbms     = "postgresql",
  server   = "localhost/testmappings",
  user     = "ohdsi_admin_user",
  password = 'admin1',
  port     = 5432,
  pathToDriver = Sys.getenv("DATABASECONNECTOR_DRIVER_POSTGRES")
)

con <- DatabaseConnector::connect(cd_admin)
```


```{r}
DatabaseConnector::disconnect(con)
```


```{r}
cdmSchema      <- "cdm_vocabs"
cdmVersion     <- "5.4"
syntheaVersion <- "2.7.0"
syntheaSchema  <- "source"
#syntheaFileLoc <- r"(C:\Users\javier\WORKSPACES\OHDSI\data\synthea_sample_data_csv_apr2020\csv)"
vocabFileLoc   <- path_to_output_omop_vocabulary_folder
```


```{r}
ETLSyntheaBuilder::CreateCDMTables(connectionDetails = cd_admin, cdmSchema = cdmSchema, cdmVersion = cdmVersion)
```

```{r}
ETLSyntheaBuilder::LoadVocabFromCsv(connectionDetails = cd_admin, cdmSchema = cdmSchema, vocabFileLoc = vocabFileLoc)
```



